On First-order Proofs without syntax

### 2021-01-05 main.tex
Fix the macros
Fix some notations

We use the notation $\vlderivation{\vlde{}{\mathcal{P}}{B}{\vlhy {A}}}$ for denoting that there is a derivation from premise $\vdash S\{A\}$ to conclusion $\vdash S\{B\}$ in system $P$ for any context $S$.

@@

We also consider the mix rule here:

\begin{center}
\begin{prooftree}
\hypo{\vdash \Gamma}
\hypo{\vdash \Delta}
\infer2[$mix$]{\vdash \Gamma, \Delta}
\end{prooftree}
\end{center}

Note that the mix rule is derivable via weakening in $\LK$.

Consider the following deep rules \cite{BT 2001}:
\begin{center}
\mbox{
\begin{prooftree}
  \hypo{\vdash S\{A \vee A\}}
  \infer1[$\cD$]{\vdash S\{A\}}
\end{prooftree}
\hspace{2cm}

@@

\begin{proposition}
\label{prop311}
Let $A$ and $B$ be two formulas such that
$\od{\odd{\odh {A}}{}{B}{\{\wD, \cD\}}}$. Then 
$\od{\odd{\odh {\PE{A}}}{}{\PE{B}}{\{\wD, \cD\}}}$.
\end{proposition}

@@

\begin{proof}
Consider the dervation $\Delta''$. If some $U_x$ (or $E_x$) is introduced
via weakening, then all the atoms it binds in $\PE{B}$ should also be introduced 
via weakening. In fact, an atom of $\PE{B}$ is introduced via weakening is 
equivalent to the fact that its corresponding vertex is not in the image of $f$. 
Since there is an edge from $U_x$ (resp. $E_x$) to all the literals it binds in the 
binding graph $\overrightarrow{G(B)}$, if one of the atoms is in the image, 
$U_x$ (resp. $E_x$) should also be in the image since $f$ is a fibration on binding graphs.

@@

We define a translation $(\cdot)^*$ inductively by:
\begin{itemize}
  \item $(E_x \vee \cdots \vee E_x)* = E_x^*$
  \item $(U_x \vee \cdots \vee U_x)* = U_x^*$
  \item structural recursion in all the other cases.
\end{itemize}

Then the derivation:
 \[\od{\odd{\odd{\odh{\PE{A}} }
  {\Delta }{A'}{\me} }
  {\Delta' }{B'}{\acD}} \]
can be translated to the derivation:
\[\od{\odd{\odh{(\PE{A})^*} }
{\Delta^* }{(B')^*}{}} \]

where $\Delta^*$ is the derivation obtained by replacing all the formulas $F$
with $F^*$.

Note that we should also introduce new rules in the derivation $\Delta^*$.
Otherwise, the rule applications cannot work in general.

To do that, we define the following rule transformation:

\begin{center}
\mbox{
\begin{prooftree}
  \hypo{Q_x^*}
  \infer1[$\acD$]{Q_x^*}
\end{prooftree}

@@

\begin{prooftree}
  \hypo{Q_x^*}
  \infer1[$=$]{Q_x^*}
\end{prooftree}

@@

\mbox{
\begin{prooftree}
  \hypo{(Q_x^* \wedge C) \vee (Q_x^* \wedge D)}
  \infer1[$\me$]{Q_x^* \wedge (C \vee D)}
\end{prooftree}

@@

\begin{prooftree}
  \hypo{(Q_x^* \wedge C) \vee (Q_x^* \wedge D)}
  \infer1[$\me'$]{Q_x^* \wedge (C \vee D)}
\end{prooftree}

@@

\end{center}
where $Q_x^*$ stands for $E_x^*$ or $U_x^*$.

$\Delta^*$ can now be transformed into a valid derivation $\Delta_1$ by using the two
transformation rules above and by applying them in a bottom-up style:
\[\od{\odd{\odh{(\PE{A})^*} }
{\Delta_1 }{(B')^*}{\acD, \me, \me'}} \]

\begin{lemma}
Every line of $\Delta_1$ is a propositional encoding (by
confounding $Q_x^*$ with $Q_x$).

\begin{proof}
We proceed by bottom-up induction in the derivation.
Clearly, $B'*$ is a propositional encoding as there is no disjunction of $Q_x$
in it.

@@

If $G_i = \forall x H_i$ for some $i$, then there will be a conjunction of $U_x^*$
and some formula which can never be eliminated by the rules $\me$, $\me'$ and
$\acD$. However, there exists no such conjunction in ${\PE{A}}^*$, which leads to a
contradiction.

@@

\begin{prooftree}
  \hypo{(Q_x^* \wedge C) \vee (Q_x^* \wedge D)}
  \infer1[$\me'$]{Q_x^* \wedge (C \vee D)}
\end{prooftree}

Suppose that $Q_X^* \wedge (C \vee D)$ = $\PE{F}$ for some $F$. It is clear
that $Q_x^* = E_x^*$ and $F = \exists x G$ with $\PE{G} = C \vee D$ for some $G$.
We distinguish two cases:
\begin{itemize}
  \item $G = \forall y H$: in this case, $(Q_x^* \wedge C) \vee (Q_x^* \wedge
D)$ has a subformula $(E_x^* \wedge U_y^*)$, which cannot be eliminated by the
	rules $\me$, $\me'$, $\acD$. It is clear that ${\PE{A}}^*$ does not
have a subformula of this form, which leads to a contradiction.
  \item $G = G_1 \vee G_2$: in this case, $(Q_x^* \wedge C) \vee (Q_x^* \wedge
  D) = \PE{((\exists x G_1) \vee (\exists x G_2))}$.
\end{itemize}

>>>>

We use the notation $\vlderivation{\vlde{}{\mathcal{P}}{B}{\vlhy {A}}}$ for
denoting that there is a derivation from premise $\vdash S\{A\}$ to conclusion
$\vdash S\{B\}$ in system $\mathcal{P}$ for any context $S$.

@@

Note that the $\mix$ rule is derivable via weakening in $\LK$.

Consider the following deep inference rules \cite{BT 2001}:
\begin{center}
\mbox{
\begin{prooftree}
  \hypo{\vdash S\{A \vee A\}}
  \infer1[$\cD$]{\vdash S\{A\}}
\end{prooftree}
\hspace{2cm}

@@

\begin{proposition}
\label{prop311}
Let $A$ and $B$ be two formulas such that
$\od{\odd{\odh {A}}{}{B}{\{\wD, \cD\}}}$. Then 
$\od{\odd{\odh {\PE{A}}}{}{\PE{B}}{\{\wD, \cD\}}}$.
\end{proposition}
\begin{proof}
  Trivial by induction.
\end{proof}

@@

\begin{proof}
Consider the dervation $\Delta''$. If some $U_x$ (or $E_x$) is introduced
via weakening, then all the atoms it binds in $\PE{B}$ should also be introduced 
via weakening. In fact, an atom of $\PE{B}$ is introduced via weakening is 
equivalent to the fact that its corresponding vertex is not in the image of $f$. 
Since there is an edge from $U_x$ (resp. $E_x$) to all the literals it binds in the 
binding graph $\overrightarrow{\Gr(B)}$, if one of the atoms is in the image, 
$U_x$ (resp. $E_x$) should also be in the image since $f$ is a fibration on binding graphs.

@@

We define a translation $(\cdot)^*$ inductively by:
\begin{itemize}
  \item $(E_x \vee \cdots \vee E_x)* = E_x$
  \item $(U_x \vee \cdots \vee U_x)* = U_x$
  \item structural recursion in all the other cases.
\end{itemize}

Then the derivation:
 \[\od{\odd{\odd{\odh{\PE{A}} }
  {\Delta }{A'}{\me} }
  {\Delta' }{\PE{B''}}{\acD}} \]
can be translated to the derivation:
\[\od{\odd{\odh{(\PE{A})^*} }
{\Delta^* }{(\PE{B''})^*}{}} \]

where $\Delta^*$ is the derivation obtained by replacing all the formulas $F$
with $F^*$ and by applying the following rule transformation:

\begin{center}
\mbox{
\begin{prooftree}
  \hypo{Q_x}
  \infer1[$\acD$]{Q_x}
\end{prooftree}

@@

\begin{prooftree}
  \hypo{Q_x}
  \infer1[$=$]{Q_x}
\end{prooftree}

@@

\mbox{
\begin{prooftree}
  \hypo{(E_x \wedge C) \vee (E_x \wedge D)}
  \infer1[$\me$]{E_x \wedge (C \vee D)}
\end{prooftree}

@@

\begin{prooftree}
  \hypo{(E_x \wedge C) \vee (E_x \wedge D)}
  \infer1[$\me'$]{E_x \wedge (C \vee D)}
\end{prooftree}

@@

\end{center}
where $Q_x$ stands for $E_x$ or $U_x$.

$\Delta^*$ can now be transformed into a valid derivation $\Delta_1$ by using the two
transformation rules above and by applying them in a bottom-up style:
\[\od{\odd{\odh{(\PE{A})^*} }
{\Delta_1 }{(\PE{B''})^*}{\acD, \me, \me'}} \]

\begin{lemma}
Every line of $\Delta_1$ is a propositional encoding.

\begin{proof}
We proceed by bottom-up induction in the derivation.
Clearly, $(\PE{B''})^*$ is a propositional encoding as there is no disjunction of $Q_x$
in it.

@@

If $G_i = \forall x H_i$ for some $i$, then there will be a conjunction of $U_x$
and some formula which can never be eliminated by the rules $\me$, $\me'$ and
$\acD$. However, there exists no such conjunction in ${\PE{A}}^*$, which leads to a
contradiction.

@@

\begin{prooftree}
  \hypo{(E_x \wedge C) \vee (E_x \wedge D)}
  \infer1[$\me'$]{E_x \wedge (C \vee D)}
\end{prooftree}

Suppose that $E_X \wedge (C \vee D)$ = $\PE{F}$ for some $F$. It is clear
that $F = \exists x G$ with $\PE{G} = C \vee D$ for some $G$.
We distinguish two cases:
\begin{itemize}
  \item $G = \forall y H$: in this case, $(E_x \wedge C) \vee (E_x \wedge
D)$ has a subformula $(E_x \wedge U_y)$, which cannot be eliminated by the
	rules $\me$, $\me'$, $\acD$. It is clear that ${\PE{A}}^*$ does not
have a subformula of this form, which leads to a contradiction.
  \item $G = G_1 \vee G_2$: in this case, $(E_x \wedge C) \vee (E_x \wedge
  D) = \PE{((\exists x G_1) \vee (\exists x G_2))}$.
\end{itemize}
